\subsection{bAbI QA Task 1: Single Supporting Fact}
\label{sec:babi_benchmark}

\paragraph{Task Description.}
The bAbI QA benchmark (Task~1) evaluates single-supporting-fact
question answering. Each sample consists of a short story describing
entity movements between locations, followed by a question of the
form \textit{''Where is Entity?''}. The correct
answer is the last location the entity moved to.

\paragraph{Test Conditions.}
The experiment evaluated 50 samples from the
\texttt{facebook/babi\_qa} dataset (subset \texttt{en-10k-qa1}).
Reasoning was performed using TransitiveResonance-validated
entity tracking: the system extracts the entity name from the
question, identifies its last movement sentence in the story,
and emits the location word. TransitiveResonance computes a
geometric validation score measuring structural plausibility
of the answer chord against the hypothesis residue.

\paragraph{Results.}
The system achieved an exact-match accuracy of 2.0000\%
across all 50 samples, with a mean partial score of
0.075.
A total of 49 samples failed exact match.
Table~\ref{tab:babi_failures} lists the first 20 failure cases.

\paragraph{Per-Entity Analysis.}
\textbf{Daniel}: 0/17 correct (0.0000\%).
\textbf{John}: 0/13 correct (0.0000\%).
\textbf{Mary}: 0/5 correct (0.0000\%).
\textbf{Sandra}: 1/15 correct (6.6667\%).

\paragraph{Per-Location Analysis.}
\textbf{bathroom}: 1/12 correct (8.3333\%).
\textbf{bedroom}: 0/8 correct (0.0000\%).
\textbf{garden}: 0/4 correct (0.0000\%).
\textbf{hallway}: 0/10 correct (0.0000\%).
\textbf{kitchen}: 0/10 correct (0.0000\%).
\textbf{office}: 0/6 correct (0.0000\%).
\begin{table}[h]
\centering
\caption{bAbI Task~1 failure cases (first 20).
Exact-match accuracy: 2.0000\%.
$N=50$, subset: \texttt{en-10k-qa1}.}
\label{tab:babi_failures}
\begin{tabular}{|r|l|l|l|}
\hline
\textbf{Idx} & \textbf{Entity} & \textbf{Expected} & \textbf{Observed} \\
\hline
0 & Mary & bathroom & Mary mov \\
\hline
1 & Daniel & hallway & Mary mo \\
\hline
2 & Daniel & hallway & arden.  \\
\hline
3 & Daniel & office &  to th \\
\hline
4 & Sandra & bathroom & is Danie \\
\hline
5 & Sandra & bathroom & the bath \\
\hline
6 & Sandra & bathroom &  Where i \\
\hline
8 & John & office & athroo \\
\hline
9 & Daniel & hallway & yed to  \\
\hline
10 & John & bathroom &  to the  \\
\hline
11 & Mary & bedroom & eyed to \\
\hline
12 & John & bathroom & neyed to \\
\hline
13 & John & bathroom & . Where  \\
\hline
14 & Sandra & bedroom & eyed to \\
\hline
15 & Daniel & bedroom & d to th \\
\hline
16 & Sandra & kitchen & d to th \\
\hline
17 & Sandra & hallway & d to th \\
\hline
18 & Sandra & hallway & d to th \\
\hline
19 & Daniel & hallway & d to th \\
\hline
20 & John & bathroom & eyed to  \\
\hline
\end{tabular}
\end{table}
