\subsection{bAbI QA Task 1: Single Supporting Fact}
\label{sec:babi_benchmark}

\paragraph{Task Description.}
The bAbI QA benchmark (Task~1) evaluates single-supporting-fact
question answering. Each sample consists of a short story describing
entity movements between named locations, followed by a question of
the form \textit{''Where is Person?''}. The correct answer is the
last location the entity moved to---requiring the system to track
an entity through a chain of movement facts without any explicit
pointer to the relevant sentence.

\paragraph{Test Conditions.}
Experiments used 50 samples from
\texttt{facebook/babi\_qa} (subset \texttt{en-10k-qa1}).
Reasoning is performed via Transitive Resonance: the entity chord is
extracted from the question, the story is scanned geometrically for
its last movement relationship, and the residue chord is decoded as
the location answer.

\paragraph{Results.}
Figure~\ref{fig:babi_trial_map} shows the per-sample Trial Outcome
Map. Each row of the left heatmap corresponds to one question;
columns show the Exact, Partial, Fuzzy, and Weighted scores on a
0--1 colour scale (viridis, dark = 0, bright = 1). The right
panel displays the weighted score per sample alongside the
overall mean (orange dashed line).

The system achieved an exact-match accuracy of 0.0\%
across all 50 samples, with a mean partial score of
0.141 and an overall weighted score of
0.064.

\paragraph{Assessment.}
Exact-match accuracy was low.  The Transitive Resonance mechanism
requires the entity's movement facts to produce a sufficiently
distinct residue chord; at this sample size the substrate geometry
may not separate location attractors reliably.

\begin{table}[htbp]
  \centering
  \caption{bAbI Task~1 failure cases (showing first 20 of
    50). $N = 50$, exact accuracy
    0.0\%.}
  \label{tab:babi_failures}
  \begin{tabular}{rlll}
    \toprule
    \textbf{Q\#} & \textbf{Entity} & \textbf{Expected} & \textbf{Observed} \\
    \midrule
    0 & Mary & \texttt{bathroom} & \texttt{bed the} \\
    1 & Daniel & \texttt{hallway} & \texttt{bed to} \\
    2 & Daniel & \texttt{hallway} & \texttt{bed to} \\
    3 & Daniel & \texttt{office} & \texttt{bathro} \\
    4 & Sandra & \texttt{bathroom} & \texttt{bed to t} \\
    5 & Sandra & \texttt{bathroom} & \texttt{bed to t} \\
    6 & Sandra & \texttt{bathroom} & \texttt{bed to t} \\
    7 & Sandra & \texttt{bathroom} & \texttt{bed to t} \\
    8 & John & \texttt{office} & \texttt{bed to} \\
    9 & Daniel & \texttt{hallway} & \texttt{bed to} \\
    10 & John & \texttt{bathroom} & \texttt{bed to t} \\
    11 & Mary & \texttt{bedroom} & \texttt{bed the} \\
    12 & John & \texttt{bathroom} & \texttt{bed to t} \\
    13 & John & \texttt{bathroom} & \texttt{bed to t} \\
    14 & Sandra & \texttt{bedroom} & \texttt{bed to} \\
    15 & Daniel & \texttt{bedroom} & \texttt{hroom.} \\
    16 & Sandra & \texttt{kitchen} & \texttt{bed to} \\
    17 & Sandra & \texttt{hallway} & \texttt{bed to} \\
    18 & Sandra & \texttt{hallway} & \texttt{bed to} \\
    19 & Daniel & \texttt{hallway} & \texttt{bed to} \\
    \bottomrule
  \end{tabular}
\end{table}
