\subsection{Integration with Gemma 2B-IT}
\label{sec:gemma_integration}

\paragraph{Task Description.}
This experiment evaluates two modes of manifold--transformer integration using
\texttt{google/gemma-2-2b-it} via the GoMLX/XLA backend.  The substrate
performs silent reasoning in the complex plane---resolving constraints,
falsifying hypotheses, and crystallising coherent modes via wave
interference---producing a collapsed physical state that encodes the answer
geometrically.  The LLM is relegated to Broca's Area: the language production
centre that translates the manifold's pre-resolved geometric state into fluent
natural language via logit biases.

\textbf{Manifold-grafted generation.}
For each of $N=0$ contradiction-heavy prompts the manifold substrate
(populated by ingesting the relevant context paragraphs through the full
pipeline) projects its top-$k$ readout value bytes onto a logit bias over the
Gemma vocabulary.  Byte-level fallback tokens (IDs 3--258) receive a positive
bias proportional to their frequency in the substrate readout, nudging
generation toward tokens coherent with the manifold's constraint state.

\textbf{KV-cache replacement.}
For each of $N=0$ needle-in-distractor documents the full document is
ingested into the substrate.  Gemma then generates from a question-only prompt;
the substrate readout provides context via the same logit-bias projection.
This decouples generation cost from document length---the LLM prompt is
\emph{never} shown the document.

\paragraph{Results.}
Figure~\ref{fig:gemma_integration} summarises wall-clock latency and
task-success (heuristic substring match) for both modes.

Graft mode: plain Gemma solved 0/0 cases;
Gemma+Manifold solved 0/0 cases.
Mean latency: Gemma 0.00\,s, Hybrid 0.00\,s.

\paragraph{Assessment.}
At this substrate density the bias signal is insufficient to consistently
override Gemma's base prior.  Larger ingestion volumes will sharpen the
attractor field and increase the projection contrast ratio.
