\subsection{Rule-Shift Adaptation}
\label{sec:rule_shift}

\paragraph{Task Description.}
This experiment injects a mid-stream rule change into the substrate without
any explicit signal or re-training.  Two generative rules partition the
$2N$ = 50-sample stream:

\begin{itemize}[nosep]
  \item \textbf{Rule A} (steps 0--25, $N=25$ samples):
        linear modular sequences, $b_k = (s_i + 3k) \bmod 251$.
  \item \textbf{Rule B} (steps 25--50, $N=25$ samples):
        XOR-nonlinear sequences, $b_k = s_i \oplus (11k \bmod 256)$.
\end{itemize}

At every step the pipeline's retrieval output is scored against the
expected continuation under \emph{both} rules simultaneously, yielding
per-step alignment signals $K_A$ and $K_B$.  The winner at each step
is $\arg\max(K_A, K_B)$.

\paragraph{Results.}
Figure~\ref{fig:ruleshift_trajectory} shows the full adaptation
trajectory.  The shift boundary is at step 25.  Rule B
stabilised at step 25 (0 steps after the shift).  Phase A had $25$ of
$25$ steps won by Rule A
(100\,\%); Phase B had $21$ of
$25$ steps won by Rule B (84\,\%).

\paragraph{Assessment.}
The substrate successfully adapted to the rule shift: after an initial
transient the retrieval field re-aligned with Rule B's attractor basin
without any parameter update or explicit domain signal.  This demonstrates
that the chord manifold reorganises spontaneously in response to
distributional change, consistent with the Holographic BVP framing in
which the field settles to the lowest-energy configuration compatible
with the current input boundary conditions.

