% ============================================================================
% ABSTRACT
% ============================================================================
\begin{abstract}
Current generative architectures treat sequence modeling as an initial value
problem: gradient-trained weights predict the next token before any future
constraint can be applied.  We reject this framing entirely and cast
generation as a \emph{boundary value problem} (BVP).  The observer clamps a
subset of symbols as fixed constraints and the remaining degrees of freedom
are resolved simultaneously via massively parallel geometric resonance search.

We instantiate this paradigm in a physics-native substrate built on three
layers:
\begin{enumerate}[nosep]
  \item An \textbf{8191-bit Value} representation over the Mersenne field
        $\operatorname{GF}(8191)$ ($2^{13}-1$), where each byte value maps to
        a deterministic prime-basis bitset across 128 blocks of 64 bits, and
        bitwise operations (\texttt{OR}, \texttt{AND}, \texttt{popcount})
        execute discrete wave interference at single-cycle throughput.
  \item A \textbf{Morton-coded Radix Trie} (Distributed Merkle Trie) that
        indexes values by their spatial keys, with sequential structure encoded
        as composable affine transformations over $\operatorname{GF}(8191)$.
  \item An \textbf{Analytical Phase} navigation layer that maps values onto a
        $S^1 \times S^1$ phase torus via their intrinsic prime-index geometry
        over $\operatorname{GF}(8191)$, enabling deterministic closure
        detection and intent steering without trained co-occurrence models.
\end{enumerate}

No loss function is minimized.  No weights are updated.  Inference is a
single-pass $\operatorname{GF}(8191)$ geometric distance scan across the
entire memory substrate, resolving the nearest rotational state via integer
arithmetic on GPU.  A two-phase hybrid retrieval pipeline (bitwise filter
followed by complex-valued PhaseDial scoring) provides precise candidate
ranking.  A self-programming loop closes the system: the cantilever BVP
solver synthesizes missing operators, successful bridges harden into native
macro-opcodes, and the interpreter dispatches them directly---reifying
execution traces into new operators.  We validate the architecture across
text classification, code generation, logical reasoning, phase-dial
generalization, image reconstruction, and cross-modal tasks, reporting both
successes and failure regimes with equal detail.
\end{abstract}
