% ============================================================================
% THE 8191-BIT CHORD REPRESENTATION
% ============================================================================
\subsection{Deterministic Prime-Basis Identity}
\label{sec:base_value}
\label{sec:value_representation}

Let $D = 8191$ denote the value dimensionality (the Mersenne prime
$2^{13} - 1$) and let $\mathcal{B} = \{0, \ldots, 255\}$ be the byte
alphabet.  Every byte value $b \in \mathcal{B}$ is mapped to a deterministic
binary vector $\mathbf{c}(b) \in \{0,1\}^{D}$ by activating exactly five bits
at positions determined by a Golomb-ruler basis projected through coprime
multipliers modulo the Mersenne prime $D = 8191$.

Let $\mathcal{G} = \{0, 1, 4, 9, 11\}$ be the 5-sparse Golomb-ruler basis and
let $\{m_1, \ldots, m_5\}$ be dynamically derived coprime multipliers
(\texttt{byteValueMultipliers}).  The activation set is:

\begin{equation}
  \mathcal{P}(b) = \bigl\{\, g_j \cdot m_j \cdot b \bmod D
                   \;\big|\; g_j \in \mathcal{G},\;
                   j = 1, \ldots, 5 \,\bigr\}
  \label{eq:base_value}
\end{equation}

\begin{equation}
  c_i(b) =
  \begin{cases}
    1 & \text{if } i \in \mathcal{P}(b), \\
    0 & \text{otherwise.}
  \end{cases}
  \label{eq:indicator}
\end{equation}

\noindent Because $D = 8191$ is a Mersenne prime ($2^{13} - 1$), every
non-zero multiplier generates a full-period permutation of the residues
$\{0, \ldots, D-1\}$, and the Golomb-ruler property guarantees that no two
basis offsets share the same pairwise difference.  Together these ensure that
distinct byte values receive distinct activation patterns (verified by
exhaustive enumeration over all $256$ values).  The resulting value is
extremely sparse: $\|\mathbf{c}(b)\|_1 = 5$ out of $D = 8191$ bits, yielding
a sparsity ratio of $5/8191 \approx 0.061\%$.

\paragraph{Why Coprime Multipliers over a Golomb Ruler.}
The choice of coprime multipliers is not arbitrary.  Multiplying by a value
$p$ and reducing modulo a prime $D = 8191$ generates a permutation of the
residues $\{0, \ldots, D-1\}$ for any $p \not\equiv 0 \pmod{D}$.
Each dynamically derived multiplier independently generates a distinct
full-period permutation.  The Golomb-ruler basis $\{0, 1, 4, 9, 11\}$
further guarantees that all pairwise differences between basis positions are
unique, maximizing the minimum Hamming distance between any two byte values
across the full 8191-bit space.

\subsection{Algebraic Operations}
\label{sec:value_algebra}

The Value supports a closed algebra of compositional operations.  Each
executes in $O(D/w)$ time where $w$ is the SIMD word width (64 for
\texttt{uint64}, yielding $\lceil D/w \rceil = 128$ operations per value):

\begin{align}
  \textbf{ValueOR:}\quad
    (\mathbf{a} \lor \mathbf{b})_i
    &= a_i \lor b_i
    &&\text{(Superposition / LCM)}
    \label{eq:value_or} \\[4pt]
  \textbf{ValueGCD:}\quad
    (\mathbf{a} \land \mathbf{b})_i
    &= a_i \land b_i
    &&\text{(Shared factors / Intersection)}
    \label{eq:value_gcd} \\[4pt]
  \textbf{ValueHole:}\quad
    H(\mathbf{t}, \mathbf{e})_i
    &= t_i \land \lnot e_i
    &&\text{(Structural Vacuum)}
    \label{eq:value_hole} \\[4pt]
  \textbf{Similarity:}\quad
    \text{sim}(\mathbf{a}, \mathbf{b})
    &= \sum_{i=0}^{D-1} a_i \cdot b_i
    = \texttt{popcount}(\mathbf{a} \land \mathbf{b})
    &&\text{(Discrete dot product)}
    \label{eq:value_sim}
\end{align}

\paragraph{Equivalence to Inner Product.}
For binary vectors $\mathbf{a}, \mathbf{b} \in \{0,1\}^D$, the real-valued
inner product $\langle \mathbf{a}, \mathbf{b} \rangle$ reduces exactly to
the popcount of their bitwise AND:
\begin{equation}
  \langle \mathbf{a}, \mathbf{b} \rangle
  = \sum_{i=0}^{D-1} a_i \, b_i
  = \bigl| \{i : a_i = 1 \land b_i = 1\} \bigr|
  = \texttt{popcount}(\mathbf{a} \mathbin{\&} \mathbf{b})
  \label{eq:inner_product}
\end{equation}
This identity is the mechanism by which single-cycle integer instructions
execute semantic similarity computation.

\subsection{Positional Binding via Circular Shift}
\label{sec:roll_left}

To encode sequential position without destroying content, we define the
\emph{RollLeft} operator as a circular permutation over bit indices:

\begin{equation}
  \text{RollLeft}(\mathbf{c},\, s)_i = c_{(i - s) \bmod D}
  \label{eq:roll_left}
\end{equation}

Because circular shifts are invertible and non-commutative with
superposition, positional binding preserves all algebraic properties while
ensuring that re-ordered sequences produce distinct values:

\begin{equation}
  \text{RollLeft}(\mathbf{c}_d, 0) \lor \text{RollLeft}(\mathbf{c}_o, 1)
  \lor \text{RollLeft}(\mathbf{c}_g, 2)
  \;\neq\;
  \text{RollLeft}(\mathbf{c}_g, 0) \lor \text{RollLeft}(\mathbf{c}_o, 1)
  \lor \text{RollLeft}(\mathbf{c}_d, 2)
  \label{eq:dog_god}
\end{equation}

\noindent That is, $\texttt{``dog''} \neq \texttt{``god''}$ in the value
representation.

\subsection{SimHash Structural Binning}
\label{sec:value_bin}

The \emph{ValueBin} operator maps any value $\mathbf{c}$ to an 8-bit
structural bin $b \in \{0, \ldots, 255\}$ via a SimHash-style random
hyperplane projection.  Given $k = 8$ fixed seed values
$\{s_1, \ldots, s_8\}$, the accumulator for plane $j$ sums signed
contributions from each active bit:

\begin{equation}
  \text{acc}_j = \sum_{i : c_i = 1}
  \begin{cases}
    +1 & \text{if } h(i, s_j) \geq 2^{63}, \\
    -1 & \text{otherwise,}
  \end{cases}
  \qquad
  h(i, s_j) = i \cdot s_j + (i \ll (j+1))
  \label{eq:simhash}
\end{equation}

The bin is then the 8-bit integer whose $j$-th bit is $[\text{acc}_j \geq 0]$.
Values with high Hamming similarity map to the same or nearby bins with
high probability, enabling the EigenMode phase table
(\Cref{sec:eigenmode_cooccurrence}) to use a fixed-size $256 \times 256$
co-occurrence matrix without per-value storage.

\subsection{FlatValue: GPU-Optimized Dense Representation}
\label{sec:flat_value}

For GPU dispatch, the sparse 8191-bit Value is converted to a \emph{FlatValue}:
a dense array of at most $D$ active prime indices $\{p_1, \ldots, p_k\}$
where $k = \|\mathbf{c}\|_1$.  This eliminates bit-twiddling thread
divergence in SIMT architectures: every thread iterates over exactly
$k$ entries rather than scanning all $\lceil D/64 \rceil = 128$ blocks.

Batched flattening across $N$ values is parallelized via a pre-warmed,
auto-scaling worker pool, yielding amortized $O(k)$ per value with
near-zero goroutine creation overhead.
