% ============================================================================
% DISCUSSION
% ============================================================================
\subsection{What Current Evidence Supports}
\label{sec:evidence_supports}

\begin{enumerate}[nosep]
  \item \textbf{Gradient-free inference:}  The system performs retrieval,
        generation, and structural reasoning without optimizing a learned loss
        surface.  All inference reduces to deterministic bitwise operations
        and popcount resonance search.
  \item \textbf{Modality-agnostic processing:}  The same byte-level tokenizer
        and Value representation processes text, images, and structured data
        through identical pipelines---no modality-specific encoders.
  \item \textbf{Non-commutative sequence encoding:}  The GF(8191) affine
        group natively encodes syntactic order through non-commutative
        composition of affine operators $f(x)=ax+b$, eliminating the
        need for learned positional embeddings.
  \item \textbf{Lossless collision-compression:}  The radix trie's Morton
        key deduplication achieves significant memory reduction while
        preserving all structural information needed for retrieval.
  \item \textbf{Hardware-native parallelism:}  The GF(8191) rotation
        resolver and PhaseDial scanner are designed for GPU warp
        architecture, with integer pre-filters and SIMD popcount
        operating at single-cycle throughput.
  \item \textbf{Self-programming:}  The system synthesizes missing affine
        operators during cantilever solving, hardens them through
        repeated success into permanent macro-opcodes, and natively
        executes the resulting learned operators---closing an
        observe-prompt-harden-execute-reify cycle without gradient
        descent.
\end{enumerate}

\subsection{Known Limitations and Failure Modes}
\label{sec:limitations}

\begin{enumerate}[nosep]
  \item \textbf{Energy selectivity:}  The current retrieval scoring is
        permissive---any candidate sharing common byte patterns scores
        highly.  Sharper gating (e.g., weighted energy functions,
        structural vacuum filtering) is needed for discriminative
        retrieval.
  \item \textbf{Propagation depth:}  Information propagation through the
        fold graph hierarchy decays with structural distance.  Deep
        multi-hop reasoning chains attenuate rapidly.
  \item \textbf{Scale competitiveness:}  The system has not been benchmarked
        against large autoregressive models on standard leaderboards.
  \item \textbf{Convergence guarantees:}  The cantilever generation loop
        converges empirically but lacks formal convergence proofs.
  \item \textbf{Sparse value coverage:}  With only 5 bits set per byte,
        an 8191-bit value is extremely sparse (${\approx}\,0.06\%$ fill
        ratio).  The capacity limits of this representation under heavy
        superposition need formal characterization.
\end{enumerate}

\subsection{Evaluation Roadmap}
\label{sec:roadmap}

\paragraph{Planned Ablation Studies.}
\begin{itemize}[nosep]
  \item Vary value width ($D \in \{4093, 8191, 16381\}$) to characterize
        the sparsity--capacity trade-off across Mersenne-prime-adjacent
        field sizes.
  \item Disable the self-programming hardening loop (no operator
        promotion) and measure the effect on cantilever prompt quality
        and convergence speed.
  \item Vary RecursiveFold depth limits to quantify the relationship
        between fold graph depth and retrieval accuracy.
\end{itemize}

\paragraph{Scaling Experiments.}
\begin{itemize}[nosep]
  \item Trie size vs.\ resolver latency: characterize the empirical
        scaling constant for 1K to 10M stored Values.
  \item Value width vs.\ retrieval accuracy: determine the minimum $D$ for
        each modality.
  \item Compression ratio vs.\ corpus size: validate the predicted
        increasing-compression-with-scale behavior from Zipfian theory
        (\Cref{sec:collision_is_compression}).
\end{itemize}

\paragraph{Novel Modality: Protein Secondary Structure.}
As a test of the architecture's genuine modality-agnosticism, we plan
experiments on protein secondary structure prediction using amino acid
sequences as raw byte input.  This domain provides a particularly
compelling test case:
\begin{enumerate}[nosep]
  \item The input alphabet (20 amino acid letters) is a strict subset
        of the byte range, requiring no architectural modification.
  \item Protein backbone geometry is governed by torsion angle pairs
        $(\phi, \psi)$ that naturally inhabit a torus $S^1 \times S^1$---
        the same manifold topology that analytical EigenMode derives
        from prime-index geometry.
  \item Secondary structure elements ($\alpha$-helices, $\beta$-sheets)
        are defined by \emph{periodic local patterns} in residue
        sequence---precisely the type of non-commutative rotational
        structure that the 8191-bit Value is designed to encode.
  \item The $O(1)$ GF(8191) rotation and popcount-based search offer an
        asymptotic advantage over molecular dynamics ($O(N^2)$ per
        timestep) for conformational space exploration.
\end{enumerate}

