% ============================================================================
% DUAL ALGEBRA: STRUCTURAL RESONANCE AND MODULAR CANCELLATION
% ============================================================================
\subsection{Two-Phase Retrieval Pipeline}
\label{sec:hybrid_pipeline}

The system resolves queries through a unified pipeline that hybridizes bitwise structural measurement with scalar algebraic cancellation over $\operatorname{GF}(8191)$:

\begin{enumerate}[nosep]
  \item \textbf{Phase 1 --- Resonance Measurement (XOR):}
        Incoming queries project a rotational trajectory into the Address Plane. The projected state is evaluated against stored anchors using bitwise XOR as a pure scalar measurement of Hamming-space distance:
        \begin{equation}
          \text{distance}(A, B) = \texttt{popcount}(A \oplus B)
          \label{eq:xor_distance}
        \end{equation}
        This measurement filters for paths experiencing constructive interference (low residue). Crucially, XOR is strictly an observable measurement gauge; it is never used to permanently compose or store state, as that would inject uncontrolled Shannon entropy.

  \item \textbf{Phase 2 --- Algebraic Cancellation:}
        Once a resonant geometric cluster is localized, logical reasoning over the localized multi-context merges resolves via scalar modular arithmetic. Because multiple facts coexist in a single cell via additive superposition $\phi_{\text{merged}} = (\phi_A + \phi_B) \pmod{8191}$, querying isolated entities involves inverse cancellation:
        \begin{equation}
          \text{result} = \phi_{\text{merged}} \cdot \phi_A^{-1} \pmod{8191} \approx \phi_B
          \label{eq:modular_inverse}
        \end{equation}
  \item \textbf{Phase 3 --- Frustration or Readout:}
        If $\text{result}$ matches a valid anchor phase ($\Delta = 1$), the wavefront locks and proceeds. Otherwise, the gap triggers \emph{frustration}, signaling the need for an extrapolated tool or a backtrack pivot.
\end{enumerate}

\subsection{Destructive Interference (Negative Constraints)}
\label{sec:negative_constraints}

Beyond positive phase-locking, the dual algebra naturally enforces negative constraints without heuristic branching rules. A negated fact (e.g., "Roy is NOT in the kitchen") is encoded as the absolute additive inverse in the field:
\begin{equation}
  \phi_{\text{neg}} = (8191 - \phi_{\text{pos}}) \pmod{8191}
  \label{eq:additive_inverse}
\end{equation}

When the exploring wavefront encounters $\phi_{\text{neg}}$, the positive baseline momentum meets the stored anti-phase. The sum algebraically collapses to zero ($\phi_{\text{pos}} + \phi_{\text{neg}} \equiv 0 \pmod{8191}$), immediately eliminating path energy and pruning the search space via pure destructive interference.
