% ============================================================================
% THE FERMAT CUBE AND ICOSAHEDRAL MANIFOLD
% ============================================================================
\subsection{Structural Layout}
\label{sec:manifold_layout}
\label{sec:manifold_architecture}

The base geometric primitive is a \emph{Fermat Cube}---a three-dimensional
array of $6 \times 4 \times 257 = 6{,}168$ Chord positions.  Each of the 6
physical sides carries 4 rotational orientations ($0^\circ$, $90^\circ$,
$180^\circ$, $270^\circ$), and each orientation holds 257 chords: 256 data
faces (one per byte value) plus a control chord at face index 256.

\begin{verbatim}
type Side [4][CubeFaces]data.Chord   // 4 rotations × 257 chords
type Cube struct {
    Sides [6]Side                     // 6 physical faces
}
\end{verbatim}

\noindent The control chord at face 256 is structurally protected: data
ingestion never writes to it.  In the Cortex (§\ref{sec:dual_architecture}),
the control chord determines edge opcodes between connected cubes.

All rotational addressing is performed via composable affine transformations
over the Galois field $\operatorname{GF}(257)$:

\begin{equation}
  f(x) = (a \cdot x + b) \pmod{257}, \qquad
  a \in \{1, \ldots, 256\},\; b \in \{0, \ldots, 256\}
  \label{eq:gf_affine}
\end{equation}

Because $257 = 2^8 + 1$ is a Fermat prime, $\operatorname{GF}(257)^*$ is
cyclic of order 256---every non-zero element generates the full
multiplicative group.  This ensures that \emph{every} byte value maps
bijectively to a unique face, and composition of rotations is closed under
the group operation (§\ref{sec:composition}).

\subsection{The Icosahedral Manifold and $A_5$ Extension}
\label{sec:icosahedral_manifold}

Each memory primitive is pre-allocated as an \texttt{IcosahedralManifold}:
5 Fermat Cubes stored contiguously with a compact orientational header:

\begin{verbatim}
type IcosahedralManifold struct {
    Header  ManifoldHeader   // 16-bit orientational state
    Cubes   [5]FermatCube    // 5 × 257 = 1,285 chord faces per cube
}
\end{verbatim}

\noindent In flat addressing mode (one cube), total state per Fermat Cube is
$6 \times 4 \times 257 \times 257 = 1{,}585{,}176$ bits.

In baseline Cubic mode (\texttt{State=0}), only \texttt{Cubes[0]} is active.

\paragraph{$A_5$ Macro-Permutations.}
Post-mitosis, macro-topology is governed by the alternating group
$A_5 \cong \text{Sym}(\text{Icosahedron})$, which consists exclusively
of \emph{even permutations} of 5 elements.  Three generator types suffice:

\begin{align}
  \textbf{3-Cycle:}\quad
    \pi_{(a\,b\,c)} &: \text{Cubes}[a] \to \text{Cubes}[b] \to
                       \text{Cubes}[c] \to \text{Cubes}[a]
  \label{eq:3cycle} \\[3pt]
  \textbf{Double Transposition:}\quad
    \pi_{(a\,b)(c\,d)} &: \text{Cubes}[a] \leftrightarrow \text{Cubes}[b],\;\;
                          \text{Cubes}[c] \leftrightarrow \text{Cubes}[d]
  \label{eq:double_trans} \\[3pt]
  \textbf{5-Cycle:}\quad
    \pi_{(a\,b\,c\,d\,e)} &: \text{Cubes}[a] \to \text{Cubes}[b] \to
                              \cdots \to \text{Cubes}[e] \to \text{Cubes}[a]
  \label{eq:5cycle}
\end{align}

Each permutation executes as an array index swap in $O(1)$ without data
copying.

\subsection{The 16-Bit Rotation Header}
\label{sec:rotation_header}

A compact \texttt{uint16} header tracks the full orientational history:

\begin{center}
\begin{tabular}{cccc}
  \toprule
  \textbf{Bit 15} & \textbf{Bits 14--9} & \textbf{Bits 8--5} & \textbf{Bits 4--0} \\
  \midrule
  State & RotState & Winding & Reserved \\
  $\{$Cubic, Ico$\}$ &
  $\in [0, 59]$ & $\bmod\; 16$ & --- \\
  \bottomrule
\end{tabular}
\end{center}

The \textbf{RotState} indexes the precomputed Unified Geodesic Matrix---a
symmetric $60 \times 60$ lookup table storing pairwise angular distances
$d(r_i, r_j) = \arccos\bigl(|q_i \cdot q_j|\bigr)$ between all $A_5$
orientations, requiring only $60 \times 60 = 3{,}600$ bytes (fits in L1
cache).  Because no floating-point quaternion math occurs during encoding,
quaternion drift is physically impossible.

\subsection{Virtual Mitosis: $O(1)$ Phase Transition}
\label{sec:mitosis}

Mitosis triggers when the global macro-density of \texttt{Cubes[0]}
reaches a precise threshold:

\begin{equation}
  \rho_0 = \frac{\text{total active bits in Cubes[0]}}
                {\text{total chord positions} \times 257} \geq 0.45
  \label{eq:mitosis_thresh}
\end{equation}

The transition is a single-cycle state flip:

\begin{enumerate}[nosep]
  \item \texttt{Cubes[0]} retains its exact saturated state.
  \item \texttt{Cubes[1..4]} remain zeroed (empty orthogonal subspaces
        ready to receive entropy).
  \item \texttt{State} flips from 0 to 1, unlocking 60 rotational states.
\end{enumerate}

\paragraph{De-Mitosis (Reversible Collapse).}
When global density across all 5 cubes drops below a sparsity threshold,
the Geodesic LUT identifies the minimum-energy $A_5$ permutation to
return \texttt{Cubes[0]} to canonical orientation, the dispersed attributes
of \texttt{Cubes[1..4]} are pooled back via majority-vote fractal
condensation, and the State reverts to 0.

\paragraph{Entropy Budget.}
In Cubic mode ($|\mathbf{O}| = 24$), the rotational register carries
$\log_2 24 \approx 4.58$ bits of structural entropy.  After mitosis
($|A_5| = 60$), this expands to $\log_2 60 \approx 5.91$ bits.  Combined
with the 4-bit winding counter ($\log_2 16 = 4$ bits), the total
orientational entropy budget is $\approx 9.91$ bits per manifold---sufficient
to distinguish $2^{9.91} \approx 961$ unique structural configurations.
