% ============================================================================
% RELATED WORK (Additional to Architecture-Specific Citations)
% ============================================================================

\paragraph{Hyperdimensional Computing and VSAs.}
The Chord representation is most directly situated within Hyperdimensional
Computing (HDC) and Vector Symbolic Architectures (VSA).  Classical VSA
frameworks use high-dimensional random vectors ($D \geq 10{,}000$) with
binding, bundling, and permutation operations.  Our 257-bit Chord provides
a far more aggressive compression, trading dimensionality for the
non-commutative rotational manifold that escapes the Shannon capacity limit
via topological entropy routing (\Cref{sec:entropy_routing}).

\paragraph{Discrete Symmetry Groups in ML.}
Equivariant neural networks exploit continuous symmetry groups (rotations,
translations) to build inductive biases.  Our use of the discrete alternating
group $A_5$ (the symmetry group of the icosahedron, order 60) is
architecturally distinct: rather than learning equivariant features, we use
$A_5$ permutations as the \emph{computational substrate itself}---syntactic
events directly manipulate geometric state through group actions.

\paragraph{Content-Addressable Memory.}
Hopfield networks, modern Hopfield networks, and transformer-based
associative memories store patterns as energy minima or attention-weighted
retrievals.  Our system differs in substrate: memory is not a learned weight
matrix but a persistent geometric index of bitwise manifolds.  New
observations reshape memory immediately via deterministic insertion.

\paragraph{Byte-Level Models.}
ByT5 and MEGABYTE process raw bytes to eliminate vocabulary failures, but
retain learned encoders and autoregressive decoding.  Our Universal Tokenizer
assigns deterministic identities without learning, and downstream inference
is parallel BestFill rather than autoregressive.

\paragraph{Non-Autoregressive Generation.}
Masked language models and diffusion-based text generators produce tokens
in parallel but struggle with output coherence.  Our BVP formulation resolves
inter-position dependencies through shared manifold topology---structural
coherence is a geometric constraint, not a decoding heuristic.

% TODO: Add citations for each paragraph once bibliography is finalized.
% TODO: Consider adding comparisons to:
%   - State-space models (S4, Mamba)
%   - Energy-based models
%   - Token Merging / dynamic sparsity approaches
