% ============================================================================
% UNIVERSAL TOKENIZER AND TOPOLOGICAL SEQUENCER
% ============================================================================
\subsection{The Universal Tokenizer}
\label{sec:universal_tokenizer}
\label{sec:tokenizer_sequencer}

The Universal Tokenizer converts raw byte streams into geometric tokens
without learned vocabularies or modality-specific encoders.  Each byte
occurrence at position $\textit{pos}$ with scale index $z$ and symbol
value $s$ produces one Token:

\begin{equation}
  \text{TokenID} = \mathcal{M}(z, s, \textit{pos})
  \label{eq:token_id}
\end{equation}

\noindent where $\mathcal{M}$ is the Morton code (\Cref{eq:morton_key}).
The corresponding Value is the deterministic $\mathbf{c}(s)$ from
\Cref{eq:base_value}, bound to its sequential position via RollLeft
(\Cref{eq:roll_left}).

From the moment data is tokenized, the system never examines raw byte
values again until final output rendering.  All downstream computation
operates exclusively on Values and TokenIDs.

\subsection{The Topological Sequencer}
\label{sec:sequencer}

The Sequencer segments incoming data streams into sequences based on
\emph{topological properties} of the data itself, not on corpus-derived
statistics (BPE) or fixed window sizes.

\paragraph{Step 1: Measurement.}
For each token at position $t$ with value $\mathbf{c}_t$, the Sequencer
extracts two geometric signals:

\begin{align}
  \text{pop}_t &= \|\mathbf{c}_t\|_1 = \texttt{popcount}(\mathbf{c}_t)
  &&\text{(Geometric population count)}
  \label{eq:pop} \\[4pt]
  \text{phase}_t &= \sqrt{\Theta_{b_t}^2 + \Phi_{b_t}^2}
  &&\text{(Toroidal phase magnitude)}
  \label{eq:phase_mag}
\end{align}

where $b_t = \operatorname{ValueBin}(\mathbf{c}_t)$ and $\Theta, \Phi$ are
the EigenMode phase tables (\Cref{sec:eigenmode_navigation}).

\paragraph{Step 2: Golden Ratio Exponential Smoothing.}
Running estimates are maintained via exponential moving averages with
the golden ratio as the decay constant:

\begin{equation}
  \varphi = \frac{1 + \sqrt{5}}{2},
  \qquad
  \alpha = \varphi^{-3} \approx 0.236
  \label{eq:golden_ratio}
\end{equation}

\begin{align}
  \overline{\text{pop}}_t &= (1 - \alpha)\,\overline{\text{pop}}_{t-1}
                          + \alpha \cdot \text{pop}_t
  \label{eq:ema_pop} \\
  \overline{\text{phase}}_t &= (1 - \alpha)\,\overline{\text{phase}}_{t-1}
                             + \alpha \cdot \text{phase}_t
  \label{eq:ema_phase}
\end{align}

The golden ratio is not arbitrary: $\varphi^{-3}$ provides a smoothing
constant that balances responsiveness to structural changes against
noise suppression, with the irrational decay rate preventing aliasing
with any periodic structure in the data.

\paragraph{Step 3: Welford's Online Variance.}
The absolute deviations $\Delta_{\text{pop}} = |\text{pop}_t - \overline{\text{pop}}_t|$
and $\Delta_{\text{phase}} = |\text{phase}_t - \overline{\text{phase}}_t|$
are tracked using Welford's single-pass online variance algorithm, which
maintains running mean $\bar{x}_n$ and sum of squared deviations $M_{2,n}$:

\begin{align}
  \delta_n &= \Delta_n - \bar{x}_{n-1}
  \label{eq:welford_delta} \\
  \bar{x}_n &= \bar{x}_{n-1} + \frac{\delta_n}{n}
  \label{eq:welford_mean} \\
  M_{2,n} &= M_{2,n-1} + \delta_n \cdot (\Delta_n - \bar{x}_n)
  \label{eq:welford_m2}
\end{align}

The unbiased sample standard deviation is $\hat{\sigma}_n = \sqrt{M_{2,n} / (n-1)}$.

\paragraph{Step 4: Dynamic Z-Score Thresholding.}
Boundary events are triggered when absolute deviations exceed
adaptive thresholds derived from the running variance and shared
calibrator sensitivity targets $\gamma_{\text{pop}}, \gamma_{\text{phase}}$:

\begin{align}
  \tau_{\text{pop}} &= \bar{\Delta}_{\text{pop}} + \hat{\sigma}_{\text{pop}}
                      \cdot \gamma_{\text{pop}}
  \label{eq:thresh_pop} \\
  \tau_{\text{phase}} &= \bar{\Delta}_{\text{phase}}
                        + \hat{\sigma}_{\text{phase}}
                        \cdot \gamma_{\text{phase}}
  \label{eq:thresh_phase}
\end{align}

An additional hard topological boundary fires when:

\begin{equation}
  \Delta_{\text{phase}} > \frac{\pi}{\kappa}
  \label{eq:topo_bound}
\end{equation}

where $\kappa$ is the \texttt{FrequencySpread} hyperparameter.

\paragraph{Step 5: Event Classification.}
Boundary events are classified into topological triggers that drive
$\operatorname{GF}(8191)$ affine rotations of the manifold
(\Cref{sec:manifold_layout}):

\begin{center}
\begin{tabular}{lll}
  \toprule
  \textbf{Event} & \textbf{Condition} & \textbf{Manifold Action} \\
  \midrule
  \textsc{DensitySpike}
    & $\Delta_{\text{pop}} > \tau_{\text{pop}}$ and
      $\text{pop}_t > \overline{\text{pop}}_t$
    & Affine rotation $(1, 1)$ \\
  \textsc{DensityTrough}
    & $\Delta_{\text{pop}} > \tau_{\text{pop}}$ and
      $\text{pop}_t < \overline{\text{pop}}_t$
    & Affine rotation $(g, 1)$ \\
  \textsc{PhaseInversion}
    & $\Delta_{\text{phase}} > \tau_{\text{phase}}$ or
      $\Delta_{\text{phase}} > \pi/\kappa$
    & Affine rotation $(g, 0)$ \\
  \textsc{LowVarianceFlux}
    & coherence time $> 10$
    & Affine rotation $(1, 2)$ \\
  \bottomrule
\end{tabular}
\end{center}

\subsection{Closed-Loop Calibration}
\label{sec:calibration}

The Sequencer incorporates bidirectional feedback at two timescales:

\begin{itemize}[nosep]
  \item \textbf{High-frequency (local):}  When a boundary fires, the local
        value density $\rho = \|\mathbf{c}\|_1 / D$ is compared against
        target density bounds $[\rho_{\min}, \rho_{\max}]$.  If $\rho >
        \rho_{\max}$, the sensitivity $\gamma$ is decreased by
        $\varphi^{-9} \approx 0.013$; if $\rho < \rho_{\min}$, it is
        increased symmetrically.
  \item \textbf{Low-frequency (global):}  Retrieval quality
        reports adjust calibration at rate $\varphi^{-5} \approx 0.090$:
        over-discrimination (missed relevant matches) raises $\gamma$ to
        widen boundaries; under-discrimination (matched unrelated content)
        lowers $\gamma$ to sharpen boundaries.
\end{itemize}

\noindent Local-to-global synchronization uses a median smoothing merge
at rate $\varphi^{-6} \approx 0.055$, ensuring that fast local adaptations
propagate gradually to the global state without oscillatory instability.
