\section{Topological Entropy Routing: Bypassing Capacity Limits in Vector Symbolic Architectures}
\label{sec:entropy_routing}

\subsection{The Shannon Limit and the ``Wall of Ones''}
A persistent vulnerability in traditional Hyperdimensional Computing (HDC) and Vector Symbolic Architectures (VSA) is the strict capacity bound imposed by Shannon entropy. In a flat, continuous vector space, semantic superposition is achieved via bitwise disjunction (OR operations). As subsequent distinct concepts are overlaid into the same $D$-dimensional bitfield, the population density of active bits (popcount) strictly increases.

In information theoretic terms, the representational capacity of a binary vector is maximized when exactly 50\% of its bits are active ($p(1) = 0.5$). Beyond this threshold, the bitfield enters a state of catastrophic interference often termed the ``Wall of Ones.'' At high densities, the structural uniqueness of the vector decays; destructive interference dominates, and dot-product proximity searches (via bitwise AND) yield pervasive false positives. 

Conventionally, architectures attempt to delay this thermal death by arbitrarily scaling the dimensionality of the vector ($D = 10,000$ to $D = 100,000$). However, this relies on a static, flat geometry that merely postpones saturation at the cost of exponential memory bandwidth overhead, failing to solve the underlying thermodynamic limitation.

\subsection{Dynamic Manifolds as Entropy Relief Valves}
Rather than treating semantic memory as a static container of arbitrary width, the proposed architecture models memory as a dynamic thermodynamic engine. We confine the atomic semantic chord to a highly compressed 257-bit vector, but geometrically map $6 \times 4 \times 257 = 6{,}168$ such chords into a three-dimensional manifold (the \textit{Fermat Cube}), addressed by $\operatorname{GF}(257)$ affine transformations.

By establishing a rigid, non-commutative rotational lattice, the architecture avoids local saturation through \textbf{Topological Entropy Routing}. Ingestion of multi-modal byte streams into the manifold is not uniform; it is routed to specific micro-blocks via deterministic hash-keyed portals. 

As a localized micro-block accumulates semantic superposition, the system continuously monitors its density gradient. When the population density approaches the Shannon maximum—specifically configured in the architecture as $\rho \ge 0.45$ (\texttt{MitosisThreshold})—the manifold reacts mechanically.

\subsection{Geometric Expansion and Virtual Mitosis}
Before the local micro-block crosses the entropy peak and loses discriminative capacity, the system triggers an affine transformation over the finite field $GF(257)$. Because affine transformations $f(p) = (ap + b) \pmod{257}$ are non-commutative, this mathematical manipulation achieves two simultaneous topological functions:
\begin{enumerate}
    \item \textbf{Sequence Encoding:} The rotational trajectory inherently encodes the grammatical or temporal sequence of the incoming data, eliminating the need for artificial positional embeddings.
    \item \textbf{Entropy Diffusion:} The highly saturated, high-entropy block is mechanically swung out of the active ingestion portal and preserved in the structural background of the manifold. Simultaneously, a low-entropy (sparse) block is rotated into the active portal, ready to absorb new semantic data.
\end{enumerate}

If the global macro-density of the Fermat Cube itself reaches the 0.45 threshold, the system executes an $O(1)$ \textit{Virtual Mitosis}. The geometry is upgraded from the chiral octahedral group ($\mathbf{O}$) to the alternating group ($A_5$), expanding the topology from a single cube to the compound of five intersecting cubes ($5 \times 6{,}168 = 30{,}840$ chord positions).

\subsection{State Space Analysis: Topology as a Capacity Multiplier}
\label{sec:state_space}

The fundamental architectural insight is that \emph{representational
capacity can be increased by adding topology rather than parameters}.
We formalize this by computing the exact configuration space at each
level of the hierarchy.

\paragraph{Level 1: The 257-Bit Chord.}
A single chord occupies $D = 257$ bits of storage and admits
$2^{257} \approx 2.31 \times 10^{77}$ possible states.  This is
the Shannon-limited micro-state.

\paragraph{Level 2: The Fermat Cube ($6 \times 4 \times 257$ positions).}
$6{,}168$ chords are arranged in a three-dimensional lattice, yielding:
\begin{equation}
  |\Omega_{\text{cube}}| = (2^{257})^{6168} = 2^{1{,}585{,}176}
  \approx 10^{477{,}166}
  \label{eq:state_cube}
\end{equation}
The memory footprint is $6{,}168 \times 64 = 394{,}752$ bytes $\approx 386$~KB.

\paragraph{Level 3: The Icosahedral Manifold (Post-Mitosis).}
Five Fermat Cubes expand the lattice to $30{,}840$ chord positions:
\begin{equation}
  |\Omega_{\text{ico}}| = (2^{257})^{30840} = 2^{7{,}925{,}880}
  \approx 10^{2{,}385{,}832}
  \label{eq:state_ico}
\end{equation}
Memory: $5 \times 394{,}752 + 8 = 1{,}973{,}768$ bytes $\approx 1.9$~MB.

\paragraph{Level 4: Rotational State Multiplier.}
The 16-bit rotation header tracks orientational history.  In $\operatorname{GF}(257)$ Affine mode, each distinct configuration can exist in $257 \times 256 = 65{,}792$
geometrically distinguishable states.  Post-mitosis with the $A_5$ working primitive, this
rotational memory space interacts with a 60-state macro-permutation.  Combined with the 4-bit winding counter
($2^4 = 16$ winding states):

\begin{equation}
  |\Omega_{\text{total}}| = 2^{7{,}925{,}880} \times 65{,}792 \times 16
  = 2^{7{,}925{,}880} \times 1{,}052{,}672
  \label{eq:state_total}
\end{equation}

Because the $GF(257)$ affine transformations are \emph{non-commutative}, the
rotational path matters---two sequences of operations that arrive at
the same final affine element via different paths produce identical
bit configurations but encode different syntactic histories in the
intermediate states.  The effective state space is therefore
\emph{path-dependent}, further multiplying the distinguishable
configurations beyond the static count.

\paragraph{Comparison with Parameter-Count Scaling.}
Table~\ref{tab:capacity_comparison} contrasts the manifold's
topology-driven capacity with conventional parameter scaling.

\begin{table}[h]
  \centering
  \caption{Memory footprint vs.\ representational state space.  The
           The Fermat Cube achieves exponentially larger state spaces per byte
           by exploiting group-theoretic multipliers rather than
           increasing parameter counts.}
  \label{tab:capacity_comparison}
  \begin{tabular}{lrrr}
    \toprule
    \textbf{Architecture}
      & \textbf{Memory}
      & \textbf{State Space}
      & \textbf{$\log_{10}$ States/KB} \\
    \midrule
    257-dim float32 embedding
      & 2 KB
      & $\sim 2^{16{,}384}$ (continuous)
      & $\sim 2{,}466$ \\
    GPT-2 ($117$M params, float16)
      & 234 MB
      & $\sim 2^{1.87 \times 10^9}$
      & $\sim 2{,}406$ \\
    \addlinespace
    Single Fermat Cube (Affine)
      & 386 KB
      & $2^{1{,}585{,}176} \times 65{,}792 \times 16$
      & $\sim 1{,}236$ \\
    Single IcosahedralManifold
      & 1.9 MB
      & $2^{7{,}925{,}880} \times 65{,}792 \times 16$
      & $\sim 1{,}236$ \\
    1M Manifolds (GPU batch)
      & 1.9 TB
      & $(2^{7{,}925{,}880})^{10^6}$
      & $\sim 1{,}236$ \\
    \bottomrule
  \end{tabular}
\end{table}

\noindent The key observation is that the states-per-KB ratio is
\emph{comparable} across architectures---because information density is
ultimately governed by Shannon's limit on the bits themselves.  What
differs is the \emph{structure} of the state space:

\begin{itemize}[nosep]
  \item In a flat embedding, all $2^{D}$ states are reachable by
        arbitrary bit flips.  Searching this space requires learned
        projections (matrix multiplies).
  \item In the manifold architecture, the state space is
        \emph{factored} into $2^{D} \times |G| \times |\text{Winding}|$
        where $|G|$ is the symmetry group order.  The group structure
        imposes a \emph{coordinate system} on the state space, making
        navigation $O(1)$ via the Geodesic LUT.
\end{itemize}

The practical consequence: 1 million IcosahedralManifolds fit in
$\sim$8.6~GB of VRAM (a single consumer GPU), yet the aggregate
state space dwarfs the parameter space of the largest language models.
Because all operations reduce to popcount and integer comparison
(no floating-point matrix multiplies), the entire memory substrate is
scanned in milliseconds.

\subsection{Implications for Scaling}
Through this mechanism, the architecture explicitly respects the Shannon
limit at the micro-state (the 257-bit chord) but circumvents it at the
macro-state by converting scalar density pressure into geometric expansion.
The system naturally diffuses semantic heat across the manifold topology,
enabling $O(1)$ ambiguity resolution over millions of manifolds natively
on consumer GPU hardware.

The analogy to physical systems is precise: protein folding achieves
astronomically large conformational spaces ($\gg 10^{100}$) from
sequences of only thousands of atoms, because \emph{spatial geometry
multiplies the configuration space exponentially}.  The Fermat Cube
exploits the same principle: chord values $\times$ Galois field permutations
$\times$ rotational paths---topology as a capacity multiplier.