\subsection{Hardware-Accelerated Symbolic Logic and Write-Optimized Storage}
\label{sec:related_hardware_accel}

The Sensorium Manifold's storage substrate---an LSM-based Radix Trie indexed by Morton-coded spatial keys---bridges two traditionally separate areas of systems research: \emph{write-optimized data structures} and \emph{GPU-native symbolic computation}.

\paragraph{Log-Structured Merge Trees and Write Amplification.}
The Log-Structured Merge Tree (LSM-Tree) \citep{oneil1996lsm} was originally designed to convert random writes into sequential I/O, dramatically improving write throughput on magnetic and flash storage. Bender et al.\ \citep{bender2007introduction} formalized the write-optimization trade-off in $B^\varepsilon$-trees, demonstrating that the ``cascade'' merge strategy achieves $O(\frac{1}{B^{1-\varepsilon}} \log_{B} N)$ amortized write cost versus $O(\log_B N)$ for B-trees, at the expense of slightly increased read amplification.

The Sensorium repurposes the LSM merge for a fundamentally different function: \emph{semantic state update}. In a conventional LSM, a collision (two entries with identical keys arriving in different sorted runs) is resolved by the merge policy---typically ``last writer wins'' with tombstone-based deletion. In the Sensorium, collision resolution \emph{is} the compression mechanism (\Cref{sec:collision_is_compression}). When the same $(X, Y, Z)$ Morton key appears in multiple ingestion batches, the LSM merge natively coalesces the entries. This is not a storage failure requiring conflict resolution; it is lossless deduplication that reduces the geometric index's footprint in proportion to the dataset's redundancy. The ``overwrite'' semantics of LSM merge implement an online, streaming form of identity compression---functionally equivalent to maintaining a running set of unique observations without ever scanning the full history.

\paragraph{Morton-Coded Keys and Cache Locality.}
The choice of Morton (Z-order) coding (\Cref{sec:morton_codes}) for the LSM sort key exploits the space-filling curve's \emph{locality-preserving} property \citep{morton1966computer, samet2006foundations}: points that are spatially adjacent in the 3D $(X, Y, Z)$ coordinate system map to nearby positions in the 1D key space. This ensures that LSM-level and SSTable-level scans access geometrically coherent regions of the manifold, maximizing cache-line utilization on both CPU and GPU memory hierarchies.

\paragraph{Metal Compute Kernels and SIMD Bitwise Operations.}
The wave-domain computation---bitwise AND/OR/XOR, popcount, and circular shift on 512-bit chords---maps directly onto the SIMD instruction sets of modern hardware:

\begin{itemize}[nosep]
  \item \textbf{Apple Metal:} The Sensorium's Metal compute shaders execute chord operations across threadgroups, with each thread processing one 512-bit chord as a \texttt{uint4} $\times$ 4 vector. Popcount is implemented via the hardware \texttt{popcount} intrinsic, achieving single-cycle throughput per 32-bit lane.
  \item \textbf{NVIDIA CUDA/Triton:} Equivalent kernels dispatch chord operations across warps, with coalesced memory access patterns ensuring full memory bandwidth utilization for the \texttt{BestFill} resonance search.
\end{itemize}

The architectural insight is that the prime-basis chord representation was designed \emph{for} this hardware: every operation required for identity comparison, similarity scoring, and superposition reduces to a fixed-length bitwise instruction with deterministic latency. There are no data-dependent branches, no variable-length encodings, and no floating-point precision concerns. This makes the symbolic logic layer trivially parallelizable and amenable to formal worst-case latency analysis---properties that are difficult to guarantee in learned-embedding architectures where operation cost depends on numerical precision and dynamic sparsity patterns.

