% ============================================================================
% NEURO-VECTOR-SYMBOLIC ARCHITECTURES AND HD COMPUTING HARDWARE
% ============================================================================
\subsection{Neuro-Vector-Symbolic Architectures and HD Computing Hardware}
\label{sec:related_nvsa}

The most direct bridge between classical neural architectures and the
prime-basis Value substrate is the emerging family of
\emph{Neuro-Vector-Symbolic Architectures} (NVSA).  Hersche et al.\
\citep{hersche2023nvsa} demonstrated that combining neural front-ends with
VSA back-ends solves Raven's Progressive Matrices---a benchmark for
abstract visual reasoning---more accurately and with orders-of-magnitude
fewer gradient steps than pure deep-learning baselines.  Their key insight
is that the VSA's algebraic closure under binding and bundling provides a
\emph{native factorization substrate}: the network learns to project
perception into a high-dimensional space where combinatorial decomposition
is a single algebraic operation rather than an iterative optimization.

The system pushes this factorization principle further in two
respects:
\begin{enumerate}[nosep]
  \item \textbf{No neural front-end:}  Where NVSA uses a learned encoder
        to project pixels into VSA space, our Universal Tokenizer assigns
        deterministic prime-basis identities directly from raw bytes.
        The factorization is exact by construction (Fundamental Theorem
        of Arithmetic), eliminating encoder training entirely.
  \item \textbf{Non-commutative extension:}  Classical VSA binding operators
        (circular convolution, Hadamard products) are commutative---they
        cannot natively distinguish ``dog bites man'' from ``man bites dog''
        without positional permutations.  The $\text{GF}(8191)$ affine group
        makes non-commutativity structural: syntactic order is algebraically
        encoded by the composition path through the affine group.
\end{enumerate}

\paragraph{HD Computing as a Hardware Paradigm.}
Rahimi et al.\ \citep{rahimi2017hd} established hyperdimensional computing
as a nanoscalable paradigm, demonstrating that HD vectors with thousands
of dimensions can perform classification, language recognition, and
biosignal processing with simple, fast-learning algorithms amenable to
3D nanometer circuit technology.  Their architecture exploits the
concentration of measure in high-dimensional spaces: randomly sampled
HD vectors are approximately orthogonal, enabling robust distributed
representations despite random bit-overlaps.

Our 8191-bit Value operates at a dimensionality ($D = 8191$) that approaches
the classical HD regime ($D \geq 4{,}000$), gaining both the concentration-of-measure
benefits of high-dimensional spaces and the exact prime-factor uniqueness of
the deterministic encoding.  The compensation mechanism for any residual
collision risk is the non-commutative affine group: where classical HD relies
on dimensionality alone for collision resistance, we additionally rely on
67 million composable affine transforms under $\text{GF}(8191)$ plus the winding
counter to separate sequences that share identical Value content.  This occupies
a novel point in the design space---deterministic algebraic separation
layered on top of high-dimensional geometric separation.

\paragraph{VSA on Emerging Hardware.}
Kleyko et al.\ \citep{kleyko2022vsa} provide a comprehensive survey of VSA
as a computing framework for emerging hardware, from in-memory computing to
neuromorphic chips.  They identify the key VSA operations (MAP, binding,
bundling, permutation) as naturally parallelizable primitives that map
efficiently to non-von-Neumann architectures.  Our Metal/CUDA/CPU kernel
dispatch layer (\Cref{sec:multi_backend}) instantiates exactly this
principle: the entire GF(8191) resolver pipeline reduces to SIMD bitwise
operations with no data-dependent branching, making it trivially portable
across any hardware that supports popcount and integer comparison.
