\subsection{Hyperdimensional Computing and Vector Symbolic Architectures}
\label{sec:related_hdc_vsa}

The computational substrate is most directly situated within the framework of \emph{Hyperdimensional Computing} (HDC) and \emph{Vector Symbolic Architectures} (VSA). In the canonical VSA formulation \citep{kanerva2009hyperdimensional}, data is represented as high-dimensional random vectors (typically $D \ge 10{,}000$), and computation is performed through three algebraic operations: \emph{binding} (element-wise multiplication or circular convolution), \emph{bundling} (element-wise addition or disjunction), and \emph{permutation} (coordinate rotation). These operations form an approximate algebra over a vector space, enabling compositional representations of structured data without learned parameters.

The \emph{Value} representation maps directly onto this framework but with critical architectural departures. Each byte identity is encoded as a sparse 8191-bit binary hypervector---the prime-basis bitset---where active bits correspond to the prime factors of a deterministic hash. In VSA terminology:

\begin{itemize}[nosep]
  \item \textbf{Binding} is implemented as \texttt{RollLeft} (circular bit-shift), which encodes positional information into the hypervector. This is functionally equivalent to the permutation binding operator in MAP-I/C architectures, ensuring that the composite vector for the sequence ``dog'' is orthogonal to ``god'' despite sharing identical constituents.
  \item \textbf{Bundling} is implemented as \texttt{ValueOR} (bitwise disjunction), which superposes multiple identities into a single bitfield. This satisfies the VSA superposition axiom: the bundled vector approximately preserves similarity to each of its components when probed via \texttt{AND}-popcount.
\end{itemize}

Gayler's foundational analysis \citep{gayler2003vector} demonstrated that VSAs provide a viable substrate for commonsense reasoning, precisely because the algebraic structure of binding and bundling supports systematic compositional generalization---a property that standard dense embeddings achieve only through expensive gradient-based training. The system inherits this guarantee: the prime-basis encoding is deterministic and compositional by construction, and the capacity of the 8191-bit value is analytically bounded by Shannon entropy rather than by learned embedding quality.

The critical innovation relative to classical HDC is the \emph{Topological Entropy Routing} mechanism (\Cref{sec:entropy_routing}). Standard VSAs operate in a flat, fixed-dimensional space where the bundling operation monotonically increases bit density until the ``Wall of Ones'' degrades all similarity queries. The system escapes this Shannon limit by embedding values into the $\text{GF}(8191)$ affine group, converting scalar density pressure into algebraic expansion. This transforms the VSA from a static container into a dynamic engine with analytically unbounded capacity.
