\subsection{Prime Factorization as Identity: G\"{o}del Numbering and Arithmetic Coding}
\label{sec:related_godel_arithmetic}

The Sensorium Manifold's use of \emph{prime-basis bitsets} as deterministic, non-colliding identity encodings is a direct computational descendant of \emph{G\"{o}del numbering}---the technique introduced in G\"{o}del's 1931 incompleteness proofs, in which every formula of a formal language is assigned a unique natural number via products of prime powers. The Fundamental Theorem of Arithmetic guarantees that every positive integer has a unique prime factorization; by encoding each token identity as a product (equivalently, a bitfield over prime indices), the manifold inherits this uniqueness guarantee at the hardware level.

\paragraph{From G\"{o}del Numbers to Prime Bitsets.}
In classical G\"{o}del encoding, a sequence of symbols $\langle s_1, s_2, \ldots, s_n \rangle$ is mapped to the integer $\prod_{i=1}^{n} p_i^{s_i}$, where $p_i$ is the $i$-th prime. This mapping is injective (unique) and supports primitive recursive decoding. The Sensorium replaces exponentiation with bit-activation: each prime index either participates (bit = 1) or does not (bit = 0) in the 257-bit value. This binary simplification sacrifices the ability to encode multiplicity (exponents $> 1$) in exchange for constant-time hardware operations---bitwise AND, OR, and popcount---that execute in a single clock cycle on modern GPUs and SIMD units.

The critical property preserved from G\"{o}del's construction is \emph{collision-freeness under composition}. When two distinct token identities are bound via the \texttt{RollLeft} operator (circular bit-shift) and then superposed via \texttt{ValueOR} (bitwise disjunction), the resulting composite value is a set-union of shifted prime indices. Because the shift is position-dependent and the primes are distinct by construction, two sequences with different orderings produce distinct composite values---a direct analogue of the injectivity of G\"{o}del numbering for sequences.

\paragraph{Collision as Compression via Arithmetic Coding.}
Witten, Neal, and Cleary's foundational work on \emph{Arithmetic Coding} \citep{witten1987arithmetic} established that optimal lossless compression maps symbols to sub-intervals of $[0, 1)$ proportional to their empirical probability, achieving entropy-rate coding without the block-alignment constraints of Huffman codes. The Sensorium's collision-compression mechanism (\Cref{sec:collision_is_compression}) implements an analogous principle in the spatial domain: when two tokens share the same $(X, Y, Z)$ Morton coordinate, their geometric index entries natively coalesce via the LSM merge operation. This is not a failure mode requiring linked-list resolution; it is \emph{lossless deduplication} that reduces the index's memory footprint in proportion to the dataset's redundancy.

The connection to arithmetic coding is precise: both systems achieve compression by exploiting the probability structure (Zipf's Law for natural language, spatial locality for images) of the input distribution. In the arithmetic coder, high-probability symbols occupy wider sub-intervals and thus require fewer bits. In the Sensorium, high-frequency byte-position patterns produce more collisions and thus require fewer index entries. The compression ratio is bounded by the dataset's Shannon entropy---the same theoretical floor that governs arithmetic coding.

\paragraph{Implications for Capacity and Scaling.}
The combination of G\"{o}del-style uniqueness (at the identity level) and arithmetic-coding-style compression (at the storage level) yields a system whose memory footprint scales with the \emph{information content} of the data rather than its raw volume. The empirical demonstration on CIFAR-10 (\Cref{sec:collision_is_compression}; 98.1\% deduplication, $52.8\times$ compression) directly validates this theoretical prediction.

