% ============================================================================
% MANIFOLD CAPACITY AND NONCOMMUTATIVE PRODUCT STRUCTURE
% ============================================================================
\subsection{Manifold Capacity and Noncommutative Product Structure}
\label{sec:related_manifold_capacity}

\paragraph{Manifold Capacity Theory.}
Chung, Lee, and Sompolinsky \citep{chung2018classification} developed a
statistical-physics theory of \emph{perceptual manifold capacity}---the
maximum number of object manifolds that can be linearly separated by a
downstream readout.  Their framework establishes that capacity depends not
on ambient dimensionality alone but on the \emph{geometry} of the manifolds:
their intrinsic dimension, radius, and center correlations.  Recent
extensions \citep{cohen2024manifoldcapacity} apply this theory across
biological neural recordings and artificial networks, introducing
\emph{Geometric measures from Correlated Manifold Capacity} (GCMC) that
analytically connect neural co-variabilities to downstream task efficiency.

The system provides a concrete, fully characterized instantiation
of this theoretical framework.  Each Value entry in the radix trie
is a discrete perceptual manifold with:
\begin{itemize}[nosep]
  \item \textbf{Intrinsic dimension:}  Bounded by the number of active
        prime factors in the constituent Values (typically 5 bits per byte
        $\times$ sequence length), spanning the 8191-bit representation
        (128 blocks $\times$ 64 bits).
  \item \textbf{Radius:}  Controlled by the ValueOR superposition density
        within each stored entry.
  \item \textbf{Center correlations:}  Determined by the $\text{GF}(8191)$ affine
        state, which algebraically separates entries that share
        byte-level content but differ in syntactic structure.
\end{itemize}

The GF(8191) rotation resolver score (\Cref{sec:five_pass}) is effectively a
linear readout over this manifold space: it computes a weighted inner
product (via popcount) that separates the query manifold from all candidates.
The achievable capacity---how many distinct memory entries the resolver can
discriminate---is therefore directly governed by the geometric properties
that Chung's theory characterizes.  Formal application of manifold capacity
bounds to the Value representation is a natural direction for future work.

\paragraph{Noncommutative $C^*$-algebra Structure.}
Hataya and Hashimoto \citep{hataya2023noncommutative} provide the formal
proof that noncommutative product structures in parameter spaces induce
``powerful effects'' in learning structured features.  Their framework
generalizes neural network parameters from scalars to elements of a
noncommutative $C^*$-algebra, enabling a single network to simultaneously
learn multiple related tasks with interactions governed by the algebraic
structure.

The $\text{GF}(8191)$ affine group that governs our value transformations is a
finite noncommutative algebra with precisely the properties that Hataya and
Hashimoto formalize:
\begin{enumerate}[nosep]
  \item \textbf{Product non-commutativity:}  $f \circ g \neq
        g \circ f$ for generic affine maps $f, g \in \text{Aff}(\text{GF}(8191))$.
        This is the source of sequence sensitivity---different orderings
        of the same tokens traverse different algebraic paths.
  \item \textbf{Closure and invertibility:}  Every affine map has
        a unique inverse (since $a \neq 0$ is invertible in $\text{GF}(8191)$),
        and the composition of any two maps is again affine.  This
        guarantees that transformations are reversible, and that resolver
        lookups are well-defined.
  \item \textbf{Scale:}  With $8190 \times 8191 = 67{,}083{,}690$ group
        elements, the affine group provides a vast algebraically
        structured state space---far richer than finite rotation groups
        of comparable computational cost.
\end{enumerate}

Where Hataya and Hashimoto's framework \emph{learns} the algebraic
structure through gradient descent, our architecture \emph{prescribes} it
through the $\text{GF}(8191)$ affine geometry.  The sequential events detected by
the Sequencer (\Cref{sec:sequencer}) are the algebraic operations; the
resolver pipeline is the readout.  The question of whether this prescribed
noncommutative structure yields the ``powerful effects'' that their theory
predicts is answered empirically by the experiments in
\Cref{sec:experiments}.
