% ============================================================================
% TOPOLOGICAL NEURAL NETWORKS AND THE EMERGING TDL FRONTIER
% ============================================================================
\subsection{Topological Neural Networks and the Symbolic$\leftrightarrow$Neuro Frontier}
\label{sec:related_topological_nn}

\paragraph{Topological Graph Neural Networks.}
Horn et al.\ \citep{horn2022togl} introduced TOGL, a layer that infuses
Graph Neural Networks with global topological information derived from
persistent homology.  Their key contribution is demonstrating that standard
GNN message-passing---which propagates information along edges---misses
structural invariants that are only visible at the \emph{topological}
level: connected components ($H_0$), loops ($H_1$), and voids ($H_2$).
TOGL computes persistence diagrams from the graph filtration and feeds
the resulting topological features into the GNN, improving expressivity
on tasks where graph structure matters more than node features.

The Sensorium Manifold addresses the same expressivity gap through a
fundamentally different mechanism.  Rather than computing topological
summaries \emph{post hoc} from a graph filtration, our architecture
\emph{embeds data into a topological space ab initio}:
\begin{itemize}[nosep]
  \item \textbf{$H_0$ (connected components):}  Analogous to the
        carrier/manifold partitioning in the PrimeField---each frozen
        manifold is a distinct connected component of the memory substrate.
  \item \textbf{$H_1$ (loops):}  The toroidal phase trajectory on the
        $S^1 \times S^1$ EigenMode torus is a native $H_1$ structure---generation
        terminates when the phase trajectory closes a loop
        (\Cref{sec:geometric_closure}).
  \item \textbf{Filtration:}  The Virtual Mitosis threshold ($\rho \geq 0.45$)
        defines a natural filtration over the manifold space: as value
        density increases, topological complexity grows from a single
        MacroCube (Cubic mode) to the full $A_5$ icosahedral manifold.
\end{itemize}

The trade-off is computational: TOGL computes persistence diagrams (which
can be expensive for large graphs) to \emph{detect} topology; our system
does not detect topology---it \emph{is} the topology.  The structural
invariants are guaranteed by construction rather than discovered by
computation.

\paragraph{Topological Deep Learning as the New Frontier.}
Papillon et al.\ \citep{papillon2024tdl} position topological deep learning
(TDL) as the next frontier for relational learning, arguing that graph-based
representations are fundamentally limited to pairwise relationships, while
real-world data exhibits higher-order interactions (simplicial complexes,
cell complexes, hypergraphs).  Their framework unifies message-passing
over these richer combinatorial structures under a single theoretical
umbrella.

The Sensorium's Value superposition naturally encodes higher-order
relationships: when $n$ bytes are superposed via \texttt{ValueOR}, the
resulting 257-bit pattern captures all $\binom{n}{k}$ subset overlaps
simultaneously through the popcount inner product.  This is not pairwise
attention---it is a holographic encoding where every subset of the input
contributes to the structural fingerprint.  The $A_5$ rotational path
then imposes a strict ordering on these higher-order relationships,
converting unstructured set-overlap into geometrically precise sequence
encoding.

\paragraph{The Symbolic$\leftrightarrow$Neuro$\leftrightarrow$Symbolic Pipeline.}
A broader trend in 2024--2025 research involves ``Symbolic $\to$ Neuro
$\to$ Symbolic'' pipelines: structured data is mapped to a vector space,
processed via geometric operations, and reconstructed as interpretable
symbols.  The Sensorium implements a maximally compressed version of this
pipeline:
\begin{enumerate}[nosep]
  \item \textbf{Symbolic $\to$ Geometric:}  Raw bytes are deterministically
        mapped to 257-bit Values and inserted into the $GF(257)$ rotational
        manifold.  No learning.
  \item \textbf{Geometric Processing:}  BestFill executes massively parallel
        popcount resonance search across the memory substrate, modulated
        by EigenMode toroidal steering.
  \item \textbf{Geometric $\to$ Symbolic:}  Retrieved manifolds are decoded
        back to byte sequences via readout, producing interpretable output.
\end{enumerate}
The entire pipeline is deterministic, invertible (via the Geodesic LUT),
and executes without gradient descent.  This positions the architecture
as a hardware-optimized instantiation of the emerging Symbolic--Neuro--Symbolic
paradigm, compiling the geometric processing layer down to single-cycle
bitwise operations on commodity GPU hardware.
