\subsection{Topological Sequencing and Boundary Analysis}
\label{sec:topological_sequencing}

For the trie-based memory substrate to correctly adapt to the linguistic or structural entropy of an incoming byte stream, the system employs a \textbf{Sequencer} that converts unstructured streams into discrete \textbf{Topological Events}.

Traditional sub-word boundaries (like BPE) rely on statically computed occurrence frequencies. The architecture replaces this with a continuous, streaming formulation of the Minimum Description Length (MDL) principle to identify natural ``hinge points'' where the information distribution drastically diverges.

\subsubsection{MDL Boundary Detection and Buffering}
The Sequencer identifies boundaries by splitting sequences where the cost of separating an active segment into two independent probability distributions optimally outperforms treating them as uniform. This is evaluated with a strict Information Criterion penalty for the addition of the structural split itself.

Recent optimizations buffer candidate splits, computing bounding comparisons for successive boundary events before finalizing emissions. The system evaluates $N$ recent candidates, validating whether trailing splits are statistically redundant. If multiple splits express similar localized distributions, the Sequencer triggers a merge, effectively collapsing fragmented structural boundaries.

\subsubsection{Structural Z-Scores and Affine Transitions}
Boundary sensitivity inherently relies on dynamic Z-scores calculated through the Standard Error of the Mean (SEM). The calculation maintains real-time variance estimates, classifying structural boundaries without assuming any fixed underlying linguistic priors. 

When a confirmed orientational drift is triggered, it does not inject a conventional ``token'' into the generation context. Instead, the Sequencer emits a $\operatorname{GF}(8191)$ affine rotation $(\Delta a, \Delta b)$ that composes with the current rotational state via $(a', b') = (a \cdot \Delta a \bmod 8191,\; a \cdot \Delta b + b \bmod 8191)$. This shifts the projection angle for all subsequent data injected into the substrate, natively mapping sequence chronology to discrete topology without positional embeddings.
